---
title: "Project Code Final"
output:
  html_document: default
  word_document: default
date: "2023-05-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
## Package Libraries
```{r}
rm(list = ls())
```

```{r}

library(readr)


library(tidyverse)
# install.packages("visdat")
library(visdat) # visualize missing values
# install.packages("caret")
library(caret)
library(ggpubr)
# install.packages("car")
library(car)
# install.packages("visdat")
library(moments)

library(dplyr)
# install.packages("MASS")
library(MASS)

library(ggplot2)

library(gridExtra)

library("DescTools")

library(rpart)
library(rpart.plot)
```

# Read Dataset
```{r}
df <- read.csv("UCI_Credit_Card.csv")
df <- as_tibble(df)
glimpse(df)
```

##  Drop `ID` column
```{r}
df <- df[,-1]
head(df)
```


# Find correlation between the target variables and independent variables
```{r}
library(corrplot)
#Find the correlation of the dataset
corplotdf <- cor(df, method = "pearson")
col_gd <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(corplotdf, method = "color", col = col_gd(200),
         type = "upper", order = "hclust",
         addCoef.col = "Black",
         tl.col = "black", tl.srt = 45, number.cex = 0.3,tl.cex = 0.4)
```

#  Rename response variable to `DEFAULT` and convert to `Yes` or `No`
```{r}
df<-df %>% 
   rename(DEFAULT = default.payment.next.month)

# Replace encoding with corresponding string for EDA, we will one-hot encode later
df <- df %>%
  mutate(DEFAULT = as.factor(ifelse(DEFAULT == "1", "Yes", "No")),
         SEX = as.factor(ifelse(SEX == "1", "Male", "Female")),
         EDUCATION = as.factor(ifelse(EDUCATION == "1", "Graduate School",
                                     ifelse(EDUCATION == "2", "University",
                                            ifelse(EDUCATION == "3", "High School",
                                                   ifelse(EDUCATION == "4", "Other", "Unknown"))))),
         MARRIAGE = as.factor(ifelse(MARRIAGE == "1", "Married",
                                     ifelse(MARRIAGE == "2", "Single", "Other")))
         )
```

## Check Missing Values
```{r}
sum(is.na(df))
vis_miss(df)
```
There is no missing values

# EDA
## Visualization

```{r}
# default and limit_balance
d0 <- ggplot(df, aes(factor(DEFAULT), (LIMIT_BAL/1000))) + geom_boxplot()
d0
```
```{r}
#default and bill payment
d1 <- ggplot(df, aes(factor(DEFAULT), (BILL_AMT1/1000))) + geom_boxplot()
d1
```
```{r}
d8 <- ggplot(df, aes(factor(DEFAULT), (PAY_AMT1))) + geom_boxplot()
d8
```

```{r}
#default and bill payment
d7 <- ggplot(df, aes(factor(DEFAULT), (PAY_0))) + geom_boxplot()
d7
```


```{r}
# SEX, limit balance education
d2 <- ggplot(df, aes(factor(SEX), (LIMIT_BAL/1000), fill=EDUCATION)) + 
  geom_boxplot() +
  xlab("Gender") + 
  ylab("BLimit(x1000 NT$)") + 
  scale_fill_brewer(palette = "Accent")
d2
```

```{r}
# Balance limits ,education and gender
d3 <- ggplot(df, aes(factor(EDUCATION), (LIMIT_BAL/1000), fill=DEFAULT)) + 
  geom_boxplot() +
  xlab("Education") + 
  ylab("BLimit(x1000 NT$)") + 
  scale_fill_brewer(palette = "Paired")
d3
```
```{r}
d3 <- ggplot(df, aes(factor(MARRIAGE), (LIMIT_BAL/1000), fill=DEFAULT)) + 
  geom_boxplot() +
  xlab("MARRIAGE") + 
  ylab("BLimit(x1000 NT$)") + 
  scale_fill_brewer(palette = "Paired")
d3
```


```{r}
# Balance limit, marraige, education
d4 <- ggplot(df, aes(factor(MARRIAGE), (LIMIT_BAL/1000), fill=EDUCATION)) + 
  geom_boxplot() +
  xlab("Marriage") + 
  ylab("BLimit(x1000 NT$)") + 
  scale_fill_brewer(palette = "Paired")
d4
```
```{r}
# Balance limits ,education and gender
d6 <- ggplot(df, aes(factor(SEX), (LIMIT_BAL/1000), fill=DEFAULT)) + 
  geom_boxplot() +
  xlab("Education") + 
  ylab("BLimit(x1000 NT$)") + 
  scale_fill_brewer(palette = "Paired")
d6
```

```{r}
d5 <- ggplot(df, aes(factor(PAY_0), (BILL_AMT1/10000), fill=DEFAULT)) + 
  geom_boxplot() +
  xlab("Repayment status in September") + 
  ylab("BILL_AMT1(x1000 NT$)") + 
  scale_fill_brewer(palette = "Paired")
d5
```
```{r}
ggplot(df, aes(x = DEFAULT, fill = DEFAULT)) +
  geom_bar() +
  labs(title = "Default Status",
       x = "Default",
       y = "Count")

```
```{r}
# install.packages("gridExtra")  # Install the package
library(gridExtra)  # Load the package

```

```{r}
graph1 <- ggplot(data=df, aes(x=BILL_AMT1,fill=DEFAULT)) + geom_histogram() +
  labs(title = "BILL_AMT1", x ="BILL_AMT1",fill = "DEFAULT") +
  scale_fill_manual(values=c("#56B4E9", "#FF9999")) +
  theme(axis.text.x = element_text(angle = 45,hjust=1))

graph2 <- ggplot(data=df, aes(x=BILL_AMT2,fill=DEFAULT)) + geom_histogram() +
  labs(title = "BILL_AMT2", x ="BILL_AMT2",fill = "DEFAULT") +
   scale_fill_manual(values=c("#56B4E9", "#FF9999"))
  theme(axis.text.x = element_text(angle = 45,hjust=1))

graph3 <- ggplot(data=df, aes(x=BILL_AMT3,fill=DEFAULT)) + geom_histogram() +
  labs(title = "BILL_AMT3", x ="BILL_AMT3",fill = "DEFAULT") +
   scale_fill_manual(values=c("#56B4E9", "#FF9999"))
  theme(axis.text.x = element_text(angle = 45,hjust=1))

graph4 <- ggplot(data=df, aes(x=BILL_AMT4,fill=DEFAULT)) + geom_histogram() +
  labs(title = "BILL_AMT4", x ="BILL_AMT4",fill = "DEFAULT") +
   scale_fill_manual(values=c("#56B4E9", "#FF9999"))
  theme(axis.text.x = element_text(angle = 45,hjust=1))

grid.arrange(graph1,graph2,graph3,graph4, ncol=2)

``` 

```{r}
graph5 <- ggplot(data=df, aes(x=PAY_AMT1,fill=DEFAULT)) + geom_histogram(binwidth = 50000) +
  labs(title = "Pay_AMT1", x ="Pay_AMT1",fill = "DEFAULT") +
  theme(axis.text.x = element_text(angle = 45,hjust=1))
graph5

graph6 <- ggplot(data=df, aes(x=PAY_AMT2,fill=DEFAULT)) + geom_histogram(binwidth =  50000) +
  labs(title = "Pay_AMT2", x ="Pay_AMT2",fill = "DEFAULT") +
  theme(axis.text.x = element_text(angle = 45,hjust=1))
  
graph7 <- ggplot(data=df, aes(x=PAY_AMT3,fill=DEFAULT)) + geom_histogram(binwidth =  50000) +
  labs(title = "Pay_AMT3", x ="Pay_AMT3",fill = "DEFAULT") +
  theme(axis.text.x = element_text(angle = 45,hjust=1))

graph8 <- ggplot(data=df, aes(x=PAY_AMT4,fill=DEFAULT)) + geom_histogram(binwidth =  50000) +
  labs(title = "Pay_AMT4", x ="Pay_AMT4",fill = "DEFAULT") +
  theme(axis.text.x = element_text(angle = 45,hjust=1))

grid.arrange(graph5,graph6,graph7,graph8,ncol=2)


```
```{r}
graph1 <- ggplot(data=df, aes(x=PAY_0,fill=DEFAULT)) + geom_histogram() +
  labs(title = "PAY_0", x ="PAY_0",fill = "DEFAULT") +
  scale_fill_manual(values=c("#56B4E9", "#FF9999")) +
  theme(axis.text.x = element_text(angle = 45,hjust=1))

graph2 <- ggplot(data=df, aes(x=PAY_2,fill=DEFAULT)) + geom_histogram() +
  labs(title = "PAY_2", x ="PAY_2",fill = "DEFAULT") +
   scale_fill_manual(values=c("#56B4E9", "#FF9999"))
  theme(axis.text.x = element_text(angle = 45,hjust=1))

graph3 <- ggplot(data=df, aes(x=PAY_3,fill=DEFAULT)) + geom_histogram() +
  labs(title = "PAY_3", x ="PAY_3",fill = "DEFAULT") +
   scale_fill_manual(values=c("#56B4E9", "#FF9999"))
  theme(axis.text.x = element_text(angle = 45,hjust=1))

graph4 <- ggplot(data=df, aes(x=PAY_4,fill=DEFAULT)) + geom_histogram() +
  labs(title = "PAY_4", x ="PAY_4",fill = "DEFAULT") +
   scale_fill_manual(values=c("#56B4E9", "#FF9999"))
  theme(axis.text.x = element_text(angle = 45,hjust=1))

graph1
grid.arrange(graph1,graph2,graph3,graph4, ncol=2)
```

```{r}
# define numeric variable
df_quant <- df %>%
  select_if(is.numeric)
```

```{r}
# install.packages("corrplot")
library(corrplot)
```


```{r}
#Find the correlation of the dataset
corplotdf <- cor(df_quant, method = "pearson")
col_gd <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(corplotdf, method = "color", col = col_gd(200),
         type = "upper", order = "hclust",
         addCoef.col = "Black",
         tl.col = "black", tl.srt = 45, number.cex = 0.5,tl.cex = 0.4)
```
```{r}
# install.packages("ggthemes")
library(ggthemes)
```

```{r}
library(dplyr)
df_sampled <- df %>% sample_n(5000)

ggplot(data = df_sampled,aes(x=AGE,y =LIMIT_BAL/1000,group=DEFAULT,color=DEFAULT,
                           shape=SEX))+
  scale_shape_manual(values=c('Male' = "O", 'Female' = "+"),labels=c("Male", "Female"))+
  scale_color_manual(values=c('No' = "grey", 'Yes' = "red"),labels=c("NO", "YES"))+
  geom_point(size=4)
```

```{r}
ggplot(data = df_sampled,aes(x=PAY_AMT1/10000,y =LIMIT_BAL/1000,group=DEFAULT,color=EDUCATION,
                           shape=DEFAULT))+
  scale_shape_manual(values=c('No' = "+", 'Yes' = "o"),labels=c("NO", "YES"))+
  scale_color_manual(values=c('Graduate School' = "blue",'High School'="yellow"),labels=c("Graduate School","High School"))+
  geom_point(size=4)
```


```{r}
p1 <- ggplot(df, aes(x = PAY_AMT1, y = BILL_AMT1, col = DEFAULT)) +
  geom_point()+ geom_smooth(method=lm,fullrange=TRUE)
p1
```



#Preprocessed

## Convert datatype to `factor`
Let's convert some int variables into factors, which indicate categorical variables in R. 
```{r}
df<-df%>%
   mutate(
          PAY_0=as.factor(PAY_0),
          PAY_2=as.factor(PAY_2),
          PAY_3=as.factor(PAY_3),
          PAY_4=as.factor(PAY_4),
          PAY_5=as.factor(PAY_5),
          PAY_6=as.factor(PAY_6)
          )
```

##Hot-one coding
```{r}
# One-Hot Encode Categorical Variables (except 'Pay_n', as they indicate ordinal values)
# `fullRank = T` == `dropfirst=T`(in python), this is to prevent multicollinearity
dmy <- dummyVars(" ~ SEX + EDUCATION + MARRIAGE", data = df, fullRank = T) 
df_transformed <- data.frame(predict(dmy, newdata = df))

glimpse(df_transformed)

# Drop SEX, EDUCATION, MARRIAGE from df
df_encoded <- df[, -c(2, 3, 4)]

# Combine df with encoded columns
df_encoded <- cbind(df_encoded, df_transformed)

df_encoded <- df_encoded%>%
   mutate(
          SEX.Male =as.factor(SEX.Male),
          EDUCATION.High.School=as.factor(EDUCATION.High.School),
          EDUCATION.Other=as.factor(EDUCATION.Other),
          EDUCATION.University =as.factor(EDUCATION.University),
          EDUCATION.Unknown =as.factor(EDUCATION.Unknown),
          MARRIAGE.Other=as.factor(MARRIAGE.Other),
          MARRIAGE.Single=as.factor(MARRIAGE.Single)
          )

glimpse(df_encoded)
```

## Skewness Correction
```{r}
# Split features into df_quant (quant features), df_cat (qual features)
df_quant <- df_encoded %>%
  select_if(is.numeric)


df_cat <- df_encoded %>% 
  select_if(is.factor)
```

## Skewness Correction
```{r}
# Split features into df_quant (quant features), df_cat (qual features)
df_quant <- df_encoded %>%
  select_if(is.numeric)


df_cat <- df_encoded %>% 
  select_if(is.factor)

```


```{r}
# log/exponential transformation for positive/negative skewness
# Define skewness correction function
skew_autotransform <- function(DF, include=NULL, exclude=NULL, plot=FALSE, threshold=1) 
  {
  # Get list of column names that should be processed based on input parameters
  if (is.null(include) & is.null(exclude)) {
    colnames <- names(DF)
  } else if (!is.null(include)) {
    colnames <- include
  } else if (!is.null(exclude)) {
    colnames <- setdiff(names(DF), exclude)
  } else {
    print('No columns to process!')
  }
  
  # Helper function that checks if all values are positive
  make_positive <- function(series) {
    minimum <- min(series)
    # If minimum is negative, offset all values by a constant to move all values to positive territory
    if (minimum <= 0) {
      series <- series + abs(minimum) + 5 # offset with a large number tailored for this dataset
    }
    return(series)
  }
  
  # Go through desired columns in DataFrame
  for (col in colnames) {
    # Get column skewness
    skew <- skewness(DF[[col]])
    transformed <- TRUE
    
    if (plot) {
      # Prep the plot of original data
      par(mfrow=c(1, 2))
      hist(DF[[col]], col="lightblue", main=paste0("Original ", col), xlab=col)
    }
    
    # If skewness is larger than threshold and positively skewed; If yes, apply appropriate transformation
    if (abs(skew) > threshold & skew > 0) {
      skewType <- 'positive'
      # Make sure all values are positive
      DF[[col]] <- make_positive(DF[[col]])
      
      # Apply log transformation 
      DF[[col]] <- log(DF[[col]])
      skew_new <- skewness(DF[[col]])
    } 
    else if (abs(skew) > threshold & skew < 0) {
      skewType <- 'negative'
      # Make sure all values are positive
      DF[[col]] <- make_positive(DF[[col]])
      
        # Apply exp transformation 
      DF[[col]] <- DF[[col]]^10
      
      skew_new <- skewness(DF[[col]])
      
    } else {
      # Flag if no transformation was performed
      transformed <- FALSE
      skew_new <- skew
    }
    
    # Compare before and after if plot is True
    if (plot) {
      cat('\n ------------------------------------------------------')     
      if (transformed) {
        cat('\n', col, 'had', skewType, 'skewness of', skew)
        cat('\n Transformation yielded skewness of', skew_new)
        hist(DF[[col]], col="salmon", main=paste0("Transformed ", col), xlab=col)
      } else {
        cat('\n NO TRANSFORMATION APPLIED FOR', col, '. Skewness =', skew)
        hist(DF[[col]], col="lightblue", main=paste0("NO TRANSFORM ", col), xlab=col)
      }
    }
  }
  
  return(DF)
}
```

```{r}
# Excluding "BILL_AMT6", as corrected to be much higher value, set threshold to be -1 to 1
df_quant_skewed <- skew_autotransform(df_quant, exclude="BILL_AMT6", plot = T, threshold = 1)
```

## Outlier Removal
```{r}
cols <- names(df_quant_skewed)

tukey_rule <- function(data, col){
  Q1 <- quantile(data[[col]], 0.25)
  Q3 <- quantile(data[[col]], 0.75)
  IQR <- Q3 - Q1
  upper_lim <- quantile(data[[col]], 0.5) + 2 * IQR
  lower_lim <- quantile(data[[col]], 0.5) - 2 * IQR
  outliers <- which(data[[col]] < lower_lim | data[[col]] >= upper_lim)
  return(outliers)
}

# Identify outliers
for (i in cols) {
  outliers_Tukey <- tukey_rule(df_quant_skewed, i)
  cat("Number of outliers in column", i, "based on Tukey's method:", length(outliers_Tukey), "\n")
}
```

```{r}
# Winsorize quant features

# Winsorizing a vector means that a predefined quantum of the smallest and/or the largest values are replaced by less extreme values. Thereby the substitute values are the most extreme retained values.
# https://www.rdocumentation.org/packages/DescTools/versions/0.99.48/topics/Winsorize

cat("Descriptive Statistics Before\n")
summary(df_quant_skewed)

df_quant_winsorized <- df_quant_skewed

for (i in cols) {
  df_quant_winsorized[, i] <- Winsorize(df_quant_winsorized[, i], probs = c(0.05, 0.95))
}

cat("Descriptive Statistics After\n")
summary(df_quant_winsorized)


for (i in cols) {
  par(mfrow=c(1, 2))
  boxplot(df_quant_skewed[,i], horizontal=TRUE, main=paste("Origin Boxplot of", i), col="lightblue")
boxplot(df_quant_winsorized[,i], horizontal=TRUE, main=paste("Winsorized Boxplot of", i), col="salmon")
}


```
```{r}
# Combine final cleaned df
df_processed <- cbind(df_quant_winsorized, df_cat)
head(df_processed)
```

```{r}
names(df_processed)
```


## Split training and testing data sets
```{r}
set.seed(1)   # set a random seed 
index <- sample(30000, 6000) # random selection of indices. (20%)
```

```{r}
test<-df_processed%>%
  filter(row_number() %in% index)

training<-df_processed%>%setdiff(test)

```

```{r}
table(training$DEFAULT)
```

```{r}
status <- ggplot(data=training, aes(x=DEFAULT,fill=DEFAULT)) + 
  geom_bar()+
  labs(title = "Unbalanced Response Variable")
status
```


### Undersample the Unbalanced Dataset
```{r}
# Undersample Dataset 
library(ROSE)
training_balanced <- ovun.sample(DEFAULT~.,data = training,
                             p=0.5,seed = 1,method = "under")$data
table(training_balanced$DEFAULT)
```

```{r}
table(test$DEFAULT)
```

```{r}
status <- ggplot(data=training_balanced, aes(x=DEFAULT,fill=DEFAULT)) + 
  geom_bar()+
  labs(title = "Balanced Response Variable")
status
```
### True Positive Rate Checking (Rational Behind Undersample)
```{r}
# Balanced Dataset
logit_model_balanced<-glm(DEFAULT~.,
                 family="binomial",               
                 data=training_balanced)                   
```

```{r}
test$logit_pred_prob_balanced<-predict(logit_model_balanced,test,type="response")
test$logit_pred_class_balanced <-ifelse(test$logit_pred_prob_balanced>0.5,"Yes","No") 
test <- test%>%
   mutate(
          logit_pred_class_balanced =as.factor(logit_pred_class_balanced)
          )
confusionMatrix(test$logit_pred_class_balanced, test$DEFAULT, positive = "Yes")
867/(524+867)
```

```{r}
# Unbalanced Dataset
logit_model_unbalanced<-glm(DEFAULT~.,  # generalized linear models
                 family="binomial",               # specifying error distribution
                 data=training)  
```

```{r}
test$logit_pred_prob_unbalanced<-predict(logit_model_unbalanced,test,type="response")
test$logit_pred_class_unbalanced<-ifelse(test$logit_pred_prob_unbalanced>0.5,"Yes","No") 
test <- test%>%
   mutate(
          logit_pred_class_unbalanced =as.factor(logit_pred_class_unbalanced)
          )
confusionMatrix(test$logit_pred_class_unbalanced, test$DEFAULT,positive = "Yes")
504/(887+504)

# The TPR has dropped by much without undersampling 
```
